---
title: "Defeo ES207_4"
author: "Shelby Defeo"
date: "4/13/2022"
output: html_document
---

1. Conduct a t-test using the “head corn” yield data shown below. Can you guess/explain where the 14:1 odds come from?
The 14:1 odds may come from an examination of the p value, wich is 0.9391. With 14:1 odds, you have a 93% probability of one outcome happening.
```{r}
# 1. Conduct a t-test using the “head corn” yield data shown below. Can you guess/explain where the 14:1 odds come from?

# Ho: The two types of seed will have no difference in mean yields.
#Ha: Kiln dried seeds will have a greater mean yield than non kild dried seeds.


#Add in data
NKD <- c(1903, 1935, 1910, 2496, 2108, 1961, 2060, 1444, 1612, 1316, 1511)
NKD

KD <- c(2009, 1915, 2011, 2463, 2180, 1925, 2122, 1482, 1542, 1443, 1535)
KD

##Testing assumptions##

#Normality
qqnorm(NKD)
qqline(NKD)

qqnorm(KD)
qqline(KD)


#Equal variance
s1_NKD <- sd(NKD)
s1_NKD

s1_KD <- sd(NKD)
s1_KD

VR_1 = s1_NKD/s1_KD
VR_1 # variance ratio; also determines whether to use Welch's or pooled. There is a large varianc

#Conduct T test
t.test(NKD, KD, alternative = c("greater"), mu = 0, paired = TRUE, var.equal = TRUE, conf.level = 0.95)
```
2. Test whether the wells downgradient of remediation are significantly lower in molybdenum than the upgradient wells. Also test whether the mean concentration has changed with remediation
The downgradient wells are not significantly lower in molybdenum because the p value of 0.06981 is greater than the alpha I chose of 0.05 so we accept the null hypothesis that there is no difference.
```{r}
#2. Test whether the wells downgradient of remediation are significantly lower in molybdenum than the upgradient wells. Also test whether the mean concentration has changed with remediation

#Hypotheses
#Ho: There is no difference in the mean concentrations of the upgradient and down gradient sites.
#Ha: The mean concentration of the downgradient wells is significantly lower than the upgradient wells. (u_DG - u_UG < 0)


#Add data
DG <- c(0.850, 0.390, 0.320, 0.300, 0.300, 0.205, 0.200, 0.200, 0.140, 0.140, 0.090, 0.046, 0.035)
DG

UG <- c(6.900, 3.200, 1.700)
UG

##Testing assumptions##

#Normality
qqnorm(DG)
qqline(DG)

qqnorm(UG)
qqline(UG)


#Equal variance
s2_UG <- sd(UG)
s2_UG

s2_DG <- sd(DG)
s2_DG

VR_2 = s2_UG/s2_DG
VR_2 # variance ratio; also determines whether to use Welch's or pooled


#are the wells downgradient of remediation are significantly lower in molybdenum than the upgradient wells?
t.test(DG, UG, alternative = c("less"), mu = 0, paired = FALSE, var.equal = FALSE, conf.level = 0.95)
 
```

3. Does the North Fork, which drains terrain with calcareous (more soluble) minerals have higher conductance than the South Fork, which drains terrain with quartz-rich (less soluble) minerals?

A. State the appropriate null and alternative hypotheses to see if conductance values are the same in the two forks.
Ho: There is no difference in mean conductance between the South Fork and North Fork.
u_NF - U_SF = 0

Ha: The North Fork has a higher mean conductance than the South Fork.
u_NF - u_SF > 0

B. Determine whether a parametric or nonparametric test should be used.
The plots are approximately normal when graphed, the variance is reasonably homogenous (0.5 < 1.1 < 2), and the observations are independent. Therefore we can use a parametric test.

C. Compute an α=0.05 test and report the results.
The p-value is 0.03373, which is less than our alpha of 0.05, so we can conclude that the North Fork does have a significantly higher conductance than the South Fork.
t: 1.9458
p-value: 0.03373
alpha: 0.05
df: 18

D. Estimate the amount by which the forks differ in conductance, regardless of the test outcome.
The mean conductance of the forks differs by 64.7 micrograms/liter.

E. Illustrate and check the results with a plot.
The plots confirm my findings.
```{r}
#3. Does the North Fork, which drains terrain with calcareous (more soluble) minerals have higher conductance than the South Fork, which drains terrain with quartz-rich (less soluble) minerals?

#Import data
library(tidyverse)
Shenandoah <- read_csv("shenandoah.csv")

NF <- Shenandoah$NORTH
NF

SF <- Shenandoah$SOUTH
SF


#B. Determine whether a parametric or nonparametric test should be used.
#Normality
qqnorm(NF)
qqline(NF)

qqnorm(SF)
qqline(SF)


#Equal variance
s3_NF <- sd(NF)
s3_NF

s3_SF <- sd(SF)
s3_SF

VR_3 = s3_NF/s3_SF
VR_3 # variance ratio; also determines whether to use Welch's or pooled


#C. Compute an α=0.05 test and report the results.
t.test(NF, SF, alternative = c("greater"), mu = 0, paired = FALSE, var.equal = TRUE, conf.level = 0.95)

#D. Estimate the amount by which the forks differ in conductance, regardless of the test outcome.
NFminusSF <- 343.9 - 279.2
NFminusSF

#E. Illustrate and check the results with a plot.
plot()
```

