---
title: "Defeo ES207_4"
author: "Shelby Defeo"
date: "4/13/2022"
output: html_document
---

1. Conduct a t-test using the “head corn” yield data shown below. Can you guess/explain where the 14:1 odds come from?
The 14:1 odds may come from an examination of the p value, which is 0.06091. With 14:1 odds, you have a 93.33% probability of one outcome happening and a 6.67% of the other outcome happening.
```{r}
# 1. Conduct a t-test using the “head corn” yield data shown below. Can you guess/explain where the 14:1 odds come from?

# Ho: The two types of seed will have no difference in mean yields.
#Ha: Kiln dried seeds will have a greater mean yield than non kild dried seeds.


#Add in data
NKD <- c(1903, 1935, 1910, 2496, 2108, 1961, 2060, 1444, 1612, 1316, 1511)
NKD

KD <- c(2009, 1915, 2011, 2463, 2180, 1925, 2122, 1482, 1542, 1443, 1535)
KD

##Testing assumptions##

#Normality
qqnorm(NKD)
qqline(NKD)

qqnorm(KD)
qqline(KD)


#Equal variance
s1_NKD <- sd(NKD)
s1_NKD

s1_KD <- sd(NKD)
s1_KD

VR_1 = s1_NKD/s1_KD
VR_1 # variance ratio; also determines whether to use Welch's or pooled. There is a large variance.

#Conduct T test
t.test(NKD, KD, mu = 0, alternative = "less", paired = TRUE, var.equal = TRUE, conf.level = 0.95)
```
2. Test whether the mean concentration has changed with remediation. Test whether the wells downgradient of remediation are significantly lower in molybdenum than the upgradient wells. 

The p-value of our two tailed is 0.01041, which is smaller than our alpha of 0.05, so we reject the null and say there is a significant difference between upgradient and downgradient values.

The p-value of our one tailed is 0.005206, which is smaller than our alpha of 0.05, then we reject the null. The downgradient values are significantly lower than the upgradient wells.
```{r}
#2. Test whether the wells downgradient of remediation are significantly lower in molybdenum than the upgradient wells (one tailed). Also test whether the mean concentration has changed with remediation (two tailed).

#Hypotheses for one tailed
#Ho: There is no difference in the mean concentrations of the upgradient and down gradient sites.
#Ha: The mean concentration of the downgradient wells is significantly lower than the upgradient wells. (u_DG - u_UG < 0)


#Add data
DG <- c(0.850, 0.390, 0.320, 0.300, 0.300, 0.205, 0.200, 0.200, 0.140, 0.140, 0.090, 0.046, 0.035)
DG

UG <- c(6.900, 3.200, 1.700)
UG

##Testing assumptions##

#Normality
qqnorm(DG)
qqline(DG)

qqnorm(UG)
qqline(UG)


#Equal variance
s2_UG <- sd(UG)
s2_UG

s2_DG <- sd(DG)
s2_DG

VR_2 = s2_UG/s2_DG
VR_2 # variance ratio; also determines whether to use Welch's or pooled

#The variance is large here so I am going to use a non parametric test to evaluate the data.Also the sample is quite small.


#test whether the mean concentration has changed with remediation
wilcox.test(UG, DG, conf.int = 0.95)

#are the wells downgradient of remediation are significantly lower in molybdenum than the upgradient wells?
wilcox.test(DG, UG, alternative = "less", conf.int = 0.95)


 
```

3. Does the North Fork, which drains terrain with calcareous (more soluble) minerals have higher conductance than the South Fork, which drains terrain with quartz-rich (less soluble) minerals?

A. State the appropriate null and alternative hypotheses to see if conductance values are the same in the two forks.
Ho: There is no difference in mean conductance between the South Fork and North Fork.
u_NF - U_SF = 0

Ha: The North Fork has a higher mean conductance than the South Fork.
u_NF - u_SF > 0

B. Determine whether a parametric or nonparametric test should be used.
The plots are approximately normal when graphed, the variance is reasonably homogenous (0.5 < 1.1 < 2), and the observations are independent. Therefore we can use a parametric test.

C. Compute an α=0.05 test and report the results.
The p-value is 0.03373, which is less than our alpha of 0.05, so we can conclude that the North Fork does have a significantly higher conductance than the South Fork.
t: 1.9458
p-value: 0.03373
alpha: 0.05
df: 18

D. Estimate the amount by which the forks differ in conductance, regardless of the test outcome.
The median conductance of the forks differs by about 65 micrograms/liter.

E. Illustrate and check the results with a plot.
The plots confirm my findings that the means are different.
```{r}
#3. Does the North Fork, which drains terrain with calcareous (more soluble) minerals have higher conductance than the South Fork, which drains terrain with quartz-rich (less soluble) minerals?

#Import data
library(tidyverse)
Shenandoah <- read_csv("shenandoah.csv")

NF <- Shenandoah$NORTH
NF

SF <- Shenandoah$SOUTH
SF


#B. Determine whether a parametric or nonparametric test should be used.
#Normality
qqnorm(NF)
qqline(NF)

qqnorm(SF)
qqline(SF)


#Equal variance
s3_NF <- sd(NF)
s3_NF

s3_SF <- sd(SF)
s3_SF

VR_3 = s3_NF/s3_SF
VR_3 # variance ratio; also determines whether to use Welch's or pooled


#C. Compute an α=0.05 test and report the results.
t.test(NF, SF, alternative = c("greater"), mu = 0, paired = FALSE, var.equal = TRUE, conf.level = 0.95)

#D. Estimate the amount by which the forks differ in conductance, regardless of the test outcome.
#Using median as measure of central tendency because the data is skewed.
hist(NF)
hist(SF)

median_NF <- median(NF)
median_NF

median_SF <- median(SF)
median_SF

compare_median <- median_NF - median_SF
compare_median


#E. Illustrate and check the results with a plot.
boxplot(UG, DG,
 names=c("Upgraident","Downgradient"),
 ylab="Concentrations")
title("Concentration by Location",cex=0.8) 


```

4. Determine whether the pH of samples taken from three sets of piezometers are identical. If not identical, which groups are different from others? Which are contaminated? Be sure to check the normality and equal variance assumptions of ANOVA if using this parametric method.
Independent: group name (classified as item)
Dependent: pH values (classified as numeric)
The pH values are not identical because the p-value of 0.00209 is less than I alpha I set of 0.05. Using the Tukey method, I can see that there is a significant difference between groups 1 and 2, as well as groups 3 and 2.
```{r}
#4. Determine whether the pH of samples taken from three sets of piezometers are identical. If not identical, which groups are different from others? Which are contaminated? Be sure to check the normality and equal variance assumptions of ANOVA if using this parametric method.
group1 <- c(7.0, 7.2, 7.5, 7.7, 8.7, 7.8)
mean(group1)

group2 <- c(6.3, 6.9, 7.0, 6.4, 6.8, 6.7)
mean(group2)

group3 <- c(8.4, 7.6, 7.5, 7.4, 9.3, 9.0)
mean(group2)

groups <- data.frame(group1, group2, group3)
groups

gathered_groups <- gather(groups, key = "group", value = "pH")
gathered_groups

#Determine whether a parametric or nonparametric test should be used.
#Normality
qqnorm(group1)
qqline(group1)

qqnorm(group2)
qqline(group2)

qqnorm(group3)
qqline(group3)

#Equal variance
s4_NF <- sd(group1)
s4_NF

s4_group2 <- sd(group2)
s4_group2

s4_group3 <- sd(group3)
s4_group3
#Approximately normal and equal variance with independent observations so I am going to use a parametric test

pH_ANOVA <- aov(pH ~ group, gathered_groups)
summary(pH_ANOVA)

TukeyHSD(pH_ANOVA, conf.level = 0.95)
plot(TukeyHSD(pH_ANOVA, conf.level = 0.95), las = 1)
#aov(dependent~ independent, data)
#create two column matrix to use ANOVA
# pH ~ group
```

5. Determine whether chloride concentrations differ among springs emanating from the three rock types. If differences occur, which rock types differ from the others? Check the assumptions of ANOVA before using it.
Chloride concentrations do differ between rock types according to an ANOVA. The p value of 0.0231 is less than the the alpha I assigned of 0.05. Using the Tukey method, I determined that grano and ephemer are significantly different from each other. We can see that because in the plot the bars do not cross zero and the p-value of 0.0351578 is less than the alpha of 0.05 I assigned. The other two combinations have p-values greater than 0.05 and they do have lines that cross zero in the plot.
```{r}
#5. Determine whether chloride concentrations differ among springs emanating from the three rock types. If differences occur, which rock types differ from the others? Check the assumptions of ANOVA before using it.
#Import data
library(tidyverse)
rocks <- read.csv("feth.csv")
rocks

chloride <- rocks$Chloride
type <- rocks$rocktype

grano <- chloride[1:18]
grano

monzon <- chloride[19:35]
monzon

ephemer <- chloride[36:50]
ephemer

#Check ANOVA assumptions
#Normality
qqnorm(grano)
qqline(grano)

qqnorm(monzon)
qqline(monzon)

qqnorm(ephemer)
qqline(ephemer)

#Equal variance
s5_grano <- sd(grano)
s5_grano

s5_monzon <- sd(monzon)
s5_monzon

s5_ephemer<- sd(ephemer)
s5_ephemer
#There is some variation from normal in this data set; I am still going to preform an ANOVA because that's what this section of the homework is about, but I recognize that assuming normal distribution for this may be a bit of a stretch.


rocks_ANOVA <- aov(chloride ~ type, rocks)
summary(rocks_ANOVA)

TukeyHSD(rocks_ANOVA, conf.level = 0.95)
plot(TukeyHSD(rocks_ANOVA, conf.level = 0.95))


```

6. Test the Corbicula data to determine whether either season or year are significant factors for the number of organisms observed. If significant effects are found, test for which levels of the factor differ from others.
Both year and season have p-values less than the alpha I assigned of 0.05 (0.000986 and 5.69e-05 respectively) so they are each significant factors.For year, 1971 - 1969 has a p-value of 0.0006494 which is less than my alpha of 0.05 so that is a significant level. The other two have p-values greater than 0.05 so they are not significant. For season, Winter - Fall (p-value: 0.0007254), Winter-Spring (p-value: 0.0006643), and Winter-Summer (p-value: 0.0000959) are significant because their p-values are less than our alpha of 0.05. The other combinations have p-values greater than 0.05 so they are not significant.

```{r}
#6. Test the Corbicula data to determine whether either season or year are significant factors for the number of organisms observed. If significant effects are found, test for which levels of the factor differ from others.
library(tidyverse)
corb <- read.csv("Corb.csv", colClasses = c("Corbicula" = "numeric", "yr" = "factor", "seas" = "factor", "grp" = "factor"))
corb


corb_ANOVA <- aov(Corbicula ~ yr * seas, corb)
summary(corb_ANOVA)

TukeyHSD(corb_ANOVA, conf.level = 0.95)
plot(TukeyHSD(corb_ANOVA, conf.level = 0.95))

```

7a. Are uranium concentrations correlated with total dissolved solids in the groundwater samples? If so, describe the strength of the relation.
There is a weak positive correlation between tds and uranium concentration in the groundwater samples, as the correlation coefficient is 0.216086. 

7b. Is the relation significant for α=0.1?
With an alpha of 0.1, the relation is significant because our p-value is 0.07945, which is less than our alpha.

```{r}
#7a. Are uranium concentrations correlated with total dissolved solids in the groundwater samples? If so, describe the strength of the relation.
library(tidyverse)
uranium_data <- read.csv("urantds2.csv")
uranium_data

uranium <- uranium_data$URANIUM

tds <- uranium_data$TDS

cor(uranium, tds)


#7b. Is the relation significant for α=0.1?
cor.test(tds, uranium, 
         alternative = "greater",
         conf.level = 0.90,
         method = c("pearson"))

```

CQ1. Using this information and dataset, determine whether at least one laboratory provided higher/lower results than the others. If so, determine which labs differ from others. Please include the following in your analysis:

A. Conduct a nonparametric test to determine if at least one laboratory is different from the others. In particular, use either the Friedman test OR the Aligned-ranks test (ART)
The p-value of 2.295e-08 is less than the alpha I assigned of 0.05. This means that at least one laboratory is different from the others.

B. With the samples as blocks, perform a two-way ANOVA without replication to decide whether at least one laboratory provided higher/lower results than the others. C. If you determine that there is a difference among the labs, Use a multiple comparison test (e.g., Tukey, Holm’s) to determine which labs differ from others E. Briefly summarize the results and conclusions of your analysis
According to my two way ANOVA without replication, there is not a significant difference between labs (p-value: 0.2770), but there is a significant difference between samples (p-value: 0.0104). This means that there are different results between the parametric and non parametric test.
```{r}
#CQ1. Using this information and dataset, determine whether at least one laboratory provided higher/lower results than the others. If so, determine which labs differ from others. Please include the following in your analysis:
library(tidyverse)
lab <- read.csv("allbac.csv")
lab

#A. Conduct a nonparametric test to determine if at least one laboratory is different from the others. In particular, use either the Friedman test OR the Aligned-ranks test (ART)
friedman.test(lab$AllBac, lab$LabNo, lab$Sample)

#B. With the samples as blocks, perform a two-way ANOVA without replication to decide whether at least one laboratory provided higher/lower results than the others. C. If you determine that there is a difference among the labs, Use a multiple comparison test (e.g., Tukey, Holm’s) to determine which labs differ from others E. Briefly summarize the results and conclusions of your analysis
lab_ANOVA <- aov(lab$AllBac ~ lab$LabNo + lab$Sample)
summary(lab_ANOVA)


#Hg.aov=(aov(Hg~Site.hg+Date.hg))
# summary(Hg.aov)

```

